Incentivizing Exploration by Heterogeneous Users

### 2018-06-05 Lemmas.tex
fixed (hopefully) L=0 issue

\begin{rlemma}{Lemma}{\ref{lem:no-incentives}}
Fix an arm $i$.
Let $s \geq \exp(2/\MinProb)$, and let $\tau_s$ be the (random)
time when arm $i$ is pulled for the \Kth{s} time.
Under \AccEU{\tau_s}{\frac{p}{4\Diam d \TieDensity}},
this pull of arm $i$ is not incentivized.
\end{rlemma}

\begin{proof}
By Lemma~\ref{lem:right-choice},
under the event \AccEU{\tau_s}{\frac{\MinProb}{4\Diam d \TieDensity}},
all agent types \AgV with
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}})
> \frac{\MinProb}{2\TieDensity}$
will pull their best arm \Best{\AgV}.

By Assmption~\ref{A1},
%Lemma~\ref{lem:sdelta},
the measure of agents (across all arms)
whose best and second-best arm differ in utility by less
than $\frac{\MinProb}{2\TieDensity}$ is at most $\frac{\MinProb}{2}$.
In particular, this bound holds for agents whose best arm is $i$.
By Assumption~\ref{A3}, at least a measure \MinProb of agents has $i$
as their best arm, and thus, at least a measure $\frac{\MinProb}{2}$
will myopically pull arm $i$.
Because $1/\log(s) < \frac{\MinProb}{2}$ for
$s \geq \exp(2/\MinProb)$, arm $i$ is not payment-eligible at time
$\tau_s$.
\end{proof}

>>>>

\begin{rlemma}{Lemma}{\ref{lem:no-incentives}}
Fix an arm $i$.
Let $s \geq \exp(2/\MinProb)$, and let $\tau_s$ be the (random)
time when arm $i$ is pulled for the \Kth{s} time.
\dkedit{Let $\hat{x} = \frac{1}{2\Diam d}
  \cdot \min(\TieCutoff, \frac{\MinProb}{2\TieDensity})$.
Under \AccEU{\tau_s}{\hat{x}},}
this pull of arm $i$ is not incentivized.
\end{rlemma}

\begin{proof}
By Lemma~\ref{lem:right-choice},
under the event \dkedit{\AccEU{\tau_s}{\hat{x}}},
all agent types \AgV with
$\AgV \cdot (\ArmV{\Best{\AgV}} - \ArmV{\Second{\AgV}})
> \dkedit{2\Diam d \hat{x}}$
will pull their best arm \Best{\AgV}.
Notice that $\dkedit{2\Diam d \hat{x}
= \min(\TieCutoff, \frac{\MinProb}{2 \TieDensity})}$

By Assmption~\ref{A1},
the measure of agents (across all arms)
whose best and second-best arm differ in utility by less
than \dkedit{$\min(\TieCutoff, \frac{\MinProb}{2 \TieDensity})$}
is at most
\dkedit{$\TieDensity \cdot \min(\TieCutoff, \frac{\MinProb}{2 \TieDensity})
\leq \frac{\MinProb}{2}$}.
In particular, this bound holds for agents whose best arm is $i$.
By Assumption~\ref{A3}, at least a measure \MinProb of agents has $i$
as their best arm, and thus, at least a measure $\frac{\MinProb}{2}$
will myopically pull arm $i$.
Because $1/\log(s) < \frac{\MinProb}{2}$ for
$s \geq \exp(2/\MinProb)$, arm $i$ is not payment-eligible at time
$\tau_s$.
\end{proof}

### 2018-06-05 algorithm_payment.tex
fixed (hopefully) L=0 issue

\begin{lemma} \label{lem:no-incentives}
Fix an arm $i$.
Let $s \geq \exp(2/\MinProb)$, and let $\tau_s$ be the (random)
time when arm $i$ is pulled for the \Kth{s} time.
Under \AccEU{\tau_s}{\frac{p}{4\Diam d \TieDensity}},
this pull of arm $i$ is not incentivized.
\end{lemma}

@@

To make this intuition precise, we set
$\delta = \frac{\MinProb}{2 \TieDensity}$,
and $x = x' = \frac{\delta}{2 \Diam d}$
(which is independent of the phase number $s$).
The cutoff point between early and late phases is now set to
$\EvenLaterPhase = \max(2, \frac{30 \sigma^3}{x^3}, \exp(\frac{2}{\MinProb}))$.
We will verify below that
$\sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{n x^2}{16 \sigma^2}}
\leq \frac{n x}{2 \sigma}$
for all $n \geq \max(2, \frac{30 \sigma^3}{x^3})$.

@@

Thus, for any $s \geq \EvenLaterPhase$,
the \Kth{s} pull of arm $i$ is incentivized  
with probability at most
$24 \ARMNUM d \cdot \exp \left(
  -\frac{1.8 s \cdot \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
\right)$.
Summing over all arms and phases $s$,
adding the at most $\ARMNUM \EvenLaterPhase$ incentivizations in
the first \EvenLaterPhase phases, 
and using that $\exp(-x) \leq 1-x/2$ for $x \leq 1$,
the expected total number of arm incentivizations is at most
\begin{align*}
\lefteqn{\ARMNUM \EvenLaterPhase
  + 24 \ARMNUM^2 d \cdot \frac{1}{1 - \exp \left(
    \frac{-1.8 \MinProb^2}{256 \TieDensity^2 \Diam^2 d^2 \sigma^2}
  \right)}}  \\
& = O\left( \max\left(\frac{\ARMNUM \TieDensity^3 \Diam^3 d^3 \sigma^3}{\MinProb^3}, N\exp\left(\frac{2}{p}\right)\right)
  + \frac{\ARMNUM^2 d^3 \TieDensity^2 \sigma^2 \Diam^2}{\MinProb^2} \right)
\; = \; O\left( N\exp\left(\frac{2}{p}\right) \right).
\end{align*}

>>>>

\begin{lemma} \label{lem:no-incentives}
Fix an arm $i$.
Let $s \geq \exp(2/\MinProb)$, and let $\tau_s$ be the (random)
time when arm $i$ is pulled for the \Kth{s} time.
\dkedit{Let $\hat{x} = \frac{1}{2\Diam d}
  \cdot \min(\TieCutoff, \frac{\MinProb}{2\TieDensity})$.
Under \AccEU{\tau_s}{\hat{x}},}
this pull of arm $i$ is not incentivized.
\end{lemma}

@@

To make this intuition precise, we set
\dkedit{$x = x' = \frac{1}{2 \Diam d} \cdot \min(\TieCutoff, \frac{\MinProb}{2 \TieDensity})$}
(which is independent of the phase number $s$).
The cutoff point between early and late phases is now set to
$\EvenLaterPhase = \max(2, \frac{30 \sigma^3}{x^3}, \exp(\frac{2}{\MinProb}))$.
We will verify below that
$\sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{n x^2}{16 \sigma^2}}
\leq \frac{n x}{2 \sigma}$
for all $n \geq \max(2, \frac{30 \sigma^3}{x^3})$.

@@

Thus, for any $s \geq \EvenLaterPhase$,
the \Kth{s} pull of arm $i$ is incentivized  
with probability at most
$24 \ARMNUM d \cdot \exp \left(
\dkedit{  - \frac{1.8 s}{256 \Diam^2 d^2 \sigma^2}
  \cdot \min(\TieCutoff, \frac{\MinProb}{\TieDensity})^2}
\right)$.
Summing over all arms and phases $s$,
adding the at most $\ARMNUM \EvenLaterPhase$ incentivizations in
the first \EvenLaterPhase phases, 
and using that $\exp(-x) \leq 1-x/2$ for $x \leq 1$,
the expected total number of arm incentivizations is at most
\begin{align*}
\lefteqn{\ARMNUM \EvenLaterPhase
  + 24 \ARMNUM^2 d \cdot \frac{1}{1 - \exp \left(
\dkedit{ - \frac{1.8}{256 \Diam^2 d^2 \sigma^2}
  \cdot \min(\TieCutoff, \frac{\MinProb}{\TieDensity})^2}
  \right)}}
\\ & = O\left( \max \left(
     \frac{\ARMNUM \TieDensity^3 \Diam^3 d^3 \sigma^3}{\MinProb^3},
\dkedit{     \frac{\ARMNUM \Diam^3 d^3 \sigma^3}{\TieCutoff^3},}
     \ARMNUM \exp \left(\frac{2}{\MinProb} \right) \right)
+ \max \left(
     \frac{\ARMNUM^2 d^3 \TieDensity^2 \sigma^2 \Diam^2}{\MinProb^2},
\dkedit{ \frac{\ARMNUM^2 d^3 \sigma^2 \Diam^2}{\TieCutoff^2}}
\right) \right)
\\ & = O\left( \ARMNUM \exp \left(\frac{2}{\MinProb}\right) \right).
\end{align*}

### 2018-06-05 problem_setting.tex
fixed (hopefully) L=0 issue

We consider a multi-armed bandit setting with \ARMNUM arms.
Arm payoffs are determined by $d$ \emph{attributes} or \emph{features};
hence, arms are identified with vectors $\ArmV{i} \in \R^d$.
The \ArmV{i} are (adversarially) fixed, and unknown to the agents
and the principal.
When arm $i$ is pulled, its current utility-relevant features are
determined as $\ArmV{i} + \NoiseV$, where \NoiseV is a mean-zero independent Gaussian%
\footnote{For simplicity of notation, we assume that the variance
  $\sigma^2$ is uniform across time steps.
In fact, the analysis extends straightforwardly to any
\pfedit{continuous}
mean-zero sub-Gaussian($\sigma^2$) noise that is independent but not necessarily identically distributed across time. \pfcomment{I used continuity below where the algorithm is defined to get rid of ties.}}
noise vector $\NoiseV \sim \Normal{\vc{0}}{\sigma^2 I_{d}}$.
Here, $I_d$ denotes the $d \times d$ identity matrix.

@@

To support the formulation of our results and the analysis,
we define the following additional notation.
We let
$\Best{\AgV} \in \argmax_i \AgV \cdot \ArmV{i}$
and
$ \Second{\AgV} \in \argmax_{i \neq \Best{\AgV}} \AgV \cdot \ArmV{i}$
denote the (indices of) the best and second-best arms for an agent
with attribute vector \AgV,
breaking ties arbitrarily (but consistently).
Notice that based on Assumption~\ref{A1} below,
  \Best{\AgV} and \Second{\AgV} are unique with probability 1.
\pfcomment{Why is \Second{\AgV} unique with probability 1?  I don't think that is implied.  Can I remove that part of the statement?}

@@

\begin{assumption}[Not too many near-ties] \label{A1}
Near-ties have vanishing probability; 
that is, there exist constants
$\TieDensity, \TieCutoff \geq \frac{\MinProb}{2 \TieDensity}$
such that 
$\AlmostTied{z} \leq \TieDensity \cdot z$ for all $z \leq \TieCutoff$.
\end{assumption}

The requirement that $\TieCutoff \geq \frac{\MinProb}{2 \TieDensity}$
is not restrictive.
Because \TieDensity and \TieCutoff are both constants,
we could simply increase \TieDensity.
Perhaps even more simply, $\MinProb \leq 1/\ARMNUM$,
while \TieCutoff is a constant,
so an alternative is to decrease \MinProb by a suitable constant factor.

Assumption~\ref{A1} implies that ties happen with probability $0$.
One special case of interest we discuss below is
$\TieDensity = 0$. This case arises when there is a cutoff
\TieCutoff such that only a measure-zero set of agents has two best arms
within utility \TieCutoff of each other.
In particular, $\TieDensity = 0$ happens whenever \AgentDist is discrete,
i.e., supported on a finite set of types,
\pfedit{and each type has a unique best arm.}

\paragraph{Roles of parameters:}
Our problem setting is characterized by many parameters.
Of these, we consider \ARMNUM, $\MinProb \leq 1/\ARMNUM$ and $T$ to be
the key parameters, while we consider
$\sigma, \Diam, d, \TieDensity, \TieCutoff$ to be constant.
We track all parameters through most of our proofs,
but report final results in terms of only the three key parameters,
except where we illustrate a specific point.
\pfcomment{We report Thm 1 with $L$ in it.  Maybe we can still leave this sentence like it is.  I'll revisit later, if someone else doesn't resolve this first.}
In particular, we consider $d$ a constant, because users
typically evaluate products (or restaurants) by a small number of
relevant features.

>>>>

We consider a multi-armed bandit setting with \ARMNUM arms.
Arm payoffs are determined by $d$ \emph{attributes} or \emph{features};
hence, arms are identified with vectors $\ArmV{i} \in \R^d$.
The \ArmV{i} are (adversarially) fixed, and unknown to the agents
and the principal.
When arm $i$ is pulled, its current utility-relevant features are
determined as $\ArmV{i} + \NoiseV$, where \NoiseV is a mean-zero independent Gaussian%
\footnote{For simplicity of notation, we assume that the variance
  $\sigma^2$ is uniform across time steps.
In fact, the analysis extends straightforwardly to any
continuous mean-zero sub-Gaussian($\sigma^2$) noise that is
independent but not necessarily identically distributed across time.
\dkcomment{Are we actually stating the result in this generality now?
  If not, then why make the algorithm use continuity, as opposed to
  that we have Gaussians? If we do make the algorithm that general,
  then we should just define everything for Gaussians here.}}
noise vector $\NoiseV \sim \Normal{\vc{0}}{\sigma^2 I_{d}}$.
Here, $I_d$ denotes the $d \times d$ identity matrix.

@@

To support the formulation of our results and the analysis,
we define the following additional notation.
We let
$\Best{\AgV} \in \argmax_i \AgV \cdot \ArmV{i}$
and
$ \Second{\AgV} \in \argmax_{i \neq \Best{\AgV}} \AgV \cdot \ArmV{i}$
denote the (indices of) the best and second-best arms for an agent
with attribute vector \AgV,
breaking ties arbitrarily (but consistently).
Notice that based on Assumption~\ref{A1} below,
  \Best{\AgV} and \Second{\AgV} are unique with probability 1.
\pfcomment{Why is \Second{\AgV} unique with probability 1?  I don't think that is implied.  Can I remove that part of the statement?}
\dkcomment{It is unique with probablity 1 over the draw of agent types
  because for z=0, Assumption~\ref{A1} says that there is a measure 0
  set of agents whose difference between the best and second-best
  option is z=0. Or am I misunderstanding something?}

@@

\begin{assumption}[Not too many near-ties] \label{A1}
Near-ties have vanishing probability; 
that is, there exist constants \dkedit{$\TieCutoff > 0, \TieDensity$} such that 
$\AlmostTied{z} \leq \TieDensity \cdot z$ for all $z \leq \TieCutoff$.
\end{assumption}

\dkdelete{
% The requirement that $\TieCutoff \geq \frac{\MinProb}{2 \TieDensity}$
%is not restrictive.
%Because \TieDensity and \TieCutoff are both constants,
%we could simply increase \TieDensity.
%Perhaps even more simply, $\MinProb \leq 1/\ARMNUM$,
%while \TieCutoff is a constant,
%so an alternative is to decrease \MinProb by a suitable constant factor.
}

Assumption~\ref{A1} implies that ties happen with probability $0$.
One special case of interest we discuss below is
$\TieDensity = 0$. This case arises when there is a cutoff
\TieCutoff such that only a measure-zero set of agents has two best arms
within utility \TieCutoff of each other.
In particular, $\TieDensity = 0$ happens whenever \AgentDist is discrete,
i.e., supported on a finite set of types,
and each type has a unique best arm.

\paragraph{Roles of parameters:}
Our problem setting is characterized by many parameters.
Of these, we consider \ARMNUM, $\MinProb \leq 1/\ARMNUM$ and $T$ to be
the key parameters, \dkedit{and also give some prominence to
\TieDensity to illustrate an interesting phenomenon.}
On the other hand, we consider
$\sigma, \Diam, d$, and \TieCutoff to be constant.
We track all parameters through most of our proofs,
but report final results in terms of only the key parameters,
except where we illustrate a specific point.
In particular, we consider $d$ a constant, because users
typically evaluate products (or restaurants, etc.) by a small number of
relevant features.
